The limits of self commitment and private paternalism \thanks{}

### 2022-11-14 donde.tex
change eqref to ref, problem of extra {

We maintain the stable unit treatment value assumption (SUTVA) throughout. This means that participant $i$'s outcomes depend only on her \emph{own} values of $Z_i$ and $D_i$, not those of any other person in the experiment: there are no \emph{spillovers}. Under this assumption, a fully general model for the potential outcomes in our experiment would take the form $Y_i(d, z)$ for $d\in \{0,1\}$ and $z \in \{0, 1, 2\}$, allowing participant $i$'s potential outcome to depend \emph{both} on the treatment she actually receives, $D_i$, and the experimental arm to which she is assigned, $Z_i$. This model is too general, however, to allow us to explore the importance of paternalism. Accordingly, we impose the following standard exclusion restriction:
\begin{equation}
Y_i(d=0, z=0) = Y_i(d=0, z=2) \equiv Y_{i0}, \quad
Y_i(d=1, z=1) = Y_i(d=1, z=2) \equiv Y_{i1}.
\label{eq:exclusion}
\end{equation}
\Eqref{eq:exclusion} says that each participant's potential outcome is the same regardless of whether she \emph{chooses} a particular treatment $D=d$ or is \emph{assigned} the same treatment. Viewing $Z_i$ as an instrumental variable (IV) for the treatment $D_i$, this nothing more than the usual IV exclusion restriction.\footnote{\cite{chamberlain2011bayesian} uses an identical assumption to develop a theory of optimal treatment choice for an individual who has access to data from an RCT.} \Eqref{eq:exclusion} is plausible in our empirical setting and has testable implications that we fail to reject: see Appendix \todo{Fill in}{???} for details. Under \eqref{eq:exclusion}, each participant's observed outcome $Y_i$ is related to her potential outcomes $(Y_{i0}, Y_{i1})$ according to 
\todo[inline]{Issac: add a section \ref{bound_tot_tut_appendix} to the appendix and drop in the results for the Huber \& Mellace stuff; Frank: follow-up and draft that section of the appendix.}
\begin{equation}
    Y_i = \mathbbm{1}(Z_i =0) Y_{i0} + \mathbbm{1}(Z_i = 1)  Y_{i1}  + \mathbbm{1}(Z_i = 2) \left[(1 - C_i) Y_{i0} + C_i Y_{i1} \right].
\label{eq:potentialOutcomes}
\end{equation}

\Eqref{eq:potentialOutcomes} is the key to understanding the results that follows. By randomly assigning $Z_i=0$ and $Z_i = 1$, our experiment identifies the marginal distributions of $Y_{i0}$ and $Y_{i1}$ in the population as a whole. By randomly assigning $Z_i=2$, it likewise identifies the share of people with $C_i = 1$ (the treated), the distribution of $Y_{i1}$ for these individuals, and the distribution of $Y_{i0}$ for those with $C_i = 0$ (the untreated). Because $Z_i$ is assigned independently of pre-treatment covariates $X_i$, our design also identifies the corresponding \emph{conditional} distributions of $Y_{i0}$ and $Y_{i1}$ given $X_i$. These distributions are the ingredients that we use below to explore the importance of paternalism.

@@

In the second step we fit a linear regression model to data from the training set using regressors constructed from the proxy functions $\psi(\cdot|\text{Training})$ and $\tau(\cdot|\text{Training})$ constructed in the first step. In particular, we estimate 
\begin{equation}
Y_i = \alpha_0 + \alpha_1 \psi_i + \beta_1 (Z_i - \mathbbm{E}[Z_i]) + \beta_2 (Z_i - \mathbbm{E}[Z_i])(\tau_i - \mathbbm{E}[\tau_i]) + \epsilon_i
\label{eq:chernreg}
\end{equation}
where $\psi_i \equiv \psi(X_i|\text{Training})$ and $\tau_i \equiv \tau(X_i|\text{Training})$.\footnote{This is a slightly simpler regression than the one proposed in equation (3.1) of \cite{chernozhukov2018generic}, which involves propensity score weights. Because the random assignment of $Z$ in our experiment does \emph{not} condition on $X$, the propensity score weights in our case are constant over $X$ and hence drop out.} As shown by \cite{chernozhukov2018generic}, the coefficients $\beta_1$ and $\beta_2$ from \eqref{eq:chernreg} identify the \emph{best linear predictor} of the conditional ATE based on $\tau(\cdot|\text{Training})$, namely
\[
\beta_1 = \mathbbm{E}[\text{ATE}(X_i)] = \text{ATE}, \quad
\beta_2 = \frac{\text{Cov}[\text{ATE}(X_i), \tau_i]}{\text{Var}(\tau_i)}.
\]
If treatment effects are homogeneous we must have $\beta_2 = 0$. Rejecting this hypothesis establishes that $\tau_i$ predicts $\text{ATE}(X_i)$ and hence that $\Delta_i$ varies. Since $\tau_i$ and $\psi_i$ do not depend on the test set, inference for the regression in \eqref{eq:chernreg} is straightforward conditional on the Training/Test split.

@@

The key insight can be read directly from \eqref{eq:potentialTreatments} and \eqref{eq:potentialOutcomes}. Viewing $Z_i$ as an instrumental variable, our design can be interpreted as a \emph{pair} of RCTs, each subject to one-sided non-compliance.\footnote{For a full derivation, see \hl{Appendix ???}.} The first of these compares $Z_i=0$ to $Z_i = 2$. For each individual with $Z_i = 0$ we have $D_i = 0$ and observe $Y_{i0}$. For those with $Z_i = 2$ we have $D_i = C_i$ and observe $(1 - C_i) Y_{i0} + C_i Y_{i1}$. This is identical to a ``randomized encouragement'' design in which treatment is only available to those who are encouraged: $Z_i = 2$. Under this interpretation, those with $C_i = 1$ are ``the compliers'' and it follows that 
\begin{equation}
\frac{\mathbbm{E}(Y_i|Z_i=2) - \mathbbm{E}(Y_i|Z_i =0)}{\mathbbm{E}(D_i|Z_i=2)-\mathbbm{E}(D_i|Z_i=0)} = 
\frac{\mathbbm{E}(Y_i|Z_i=2) - \mathbbm{E}(Y_i|Z_i =0)}{\mathbbm{E}(D_i|Z_i=2)} = \mathbbm{E}(Y_{i1} - Y_{i0}|C_i = 1)
\label{eq:ToT}
\end{equation}
since $\mathbbm{E}(D_i|Z_i=0)=0$ by \eqref{eq:potentialTreatments}.

A closely related if slightly less intuitive argument can be used to construct a Wald estimand that identifies the TuT. Here we consider $Z_i = 1$ to be the ``encouragement'' and compare the outcomes for these individuals to those with $Z_i = 2$. If $Z_i = 1$ then $D_i = 1$ and we observe $Y_{i1}$; If $Z_i = 2$ then $D_i = C_i$ and we observe $(1 - C_i) Y_{i0} + C_i Y_{i1}$. Again, we can view this as an experiment with one-sided non-compliance, but now the situation is reversed. Everyone with $Z_i = 1$ is treated, but some people with $Z_i = 2$ are ``always-takers'' who obtain the treatment ($D_i = 1$) despite having been allocated to the ``control'' arm $Z_i=2$. Under this interpretation, the ``compliers'' are those with $C_i = 0$: when $Z_i=1$ they take the treatment, and when $Z_i=2$, they do not. Thus, 
\begin{equation}
\frac{\mathbbm{E}(Y_i|Z_i=1) - \mathbbm{E}(Y_i|Z_i =2)}{\mathbbm{E}(D_i|Z_i=1)-\mathbbm{E}(D_i|Z_i=2)} = 
\frac{\mathbbm{E}(Y_i|Z_i=1) - \mathbbm{E}(Y_i|Z_i =0)}{1 - \mathbbm{E}(D_i|Z_i=2)} = \mathbbm{E}(Y_{i1} - Y_{i0} | C_i = 0)
\label{eq:TuT}
\end{equation}
since $\mathbbm{E}(D_i|Z_i = 1)= 1$ by \eqref{eq:potentialTreatments}.

\Eqref{eq:ToT} and \eqref{eq:TuT} are useful for understanding why our design identifies the ToT and TuT, but they are less convenient for estimation and inference. In Appendix \todo{Frank: fill in}{???}, we show that
\begin{align}
\label{eq:ToTreg}
Y_i &= \mathbbm{E}(Y_{i0}) + (\text{ATE}) \mathbbm{1}(Z_i = 1) + (\text{ToT}) \left[\mathbbm{1}(Z_i = 2) \times D_i\right] + U_i \\
\label{eq:TuTreg}
Y_i &= \mathbbm{E}(Y_{i1}) + (\text{ATE}) \left[ -\mathbbm{1}(Z_i = 0)\right] + (\text{TuT}) \left[ -\mathbbm{1}(Z_i = 2) \times (1 - D_i)\right] + V_i 
\end{align}
where the error terms $U_i$ and $V_i$ satisfy 
\[
\mathbbm{E}(V_i) = \mathbbm{E}(U_i) = \mathbbm{E}[U_i \mathbbm{1}(Z_i = z)] = \mathbbm{E}[V_i \mathbbm{1}(Z_i = z)] = 0, \quad z = 0, 1, 2.
\]
It follows that a pair of just-identified, linear instrumental variables regressions can be used to estimate and carry out inference for each of our quantities of interest. To identify the ATE and ToT, run IV on \eqref{eq:ToTreg} with instruments $\mathbbm{1}(Z_i = 1)$ and $\mathbbm{1}(Z_i=2)$. An F-test based on this regression can then be used to test the restriction $\text{ATE} = \text{ToT}$, and the usual IV output can be used to carry out inference for the ToT. No programming is required. Similarly, to identify the ATE and TuT run IV on \ref{eq:TuTreg} with instruments $\mathbbm{1}(Z_i = 0)$ and $\mathbbm{1}(Z_i = 2)$. An F-test based on this regression can then be used to test the restriction $\text{ATE} = \text{TuT}$, and the usual IV output can be used to carry out inference for the TuT. Again, no programming or special software is required.

@@

%Define the shorthand
%\[
%p_0 \equiv \mathbbm{P}(Z_i=0), \quad p_1 \equiv \mathbbm{P}(Z_i = 1), \quad q_0 \equiv \mathbbm{E}[\mathbbm{1}(Z_i = 2)(1 - D_i)], \quad q_1 \equiv \mathbbm{E}[\mathbbm{1}(Z_i = 2) D_i].
%\]
%As shown in Appendix \todo{Frank: fill in}{???}, the asymptotic variance of $(\text{Tot} - \text{TuT})$ is given by
%\[
%\text{AVAR}(\text{ToT} - \text{TuT}) = \boldsymbol{\omega}' \boldsymbol{\Sigma} \boldsymbol{\omega}, \quad
%\boldsymbol{\omega}' \equiv \displaystyle \begin{bmatrix}
%-\frac{1}{q_1} & \frac{1}{q_0} & \left(\frac{p_1 - 1}{p_0q_1} - \frac{1}{q_0}\right) & -\frac{1}{q_1} - \frac{1 - p_0}{p_1 q_0}
%\end{bmatrix}
%\]
%where $\boldsymbol{\Sigma}$ is a covariance matrix that can be estimated from the residuals of \eqref{eq:ToTreg} and \eqref{eq:TuTreg}. Under iid observations, 
%
%\todo[inline]{Explain two leading forms that $\Sigma$ can take: heterosked and cluster}

>>>>

We maintain the stable unit treatment value assumption (SUTVA) throughout. This means that participant $i$'s outcomes depend only on her \emph{own} values of $Z_i$ and $D_i$, not those of any other person in the experiment: there are no \emph{spillovers}. Under this assumption, a fully general model for the potential outcomes in our experiment would take the form $Y_i(d, z)$ for $d\in \{0,1\}$ and $z \in \{0, 1, 2\}$, allowing participant $i$'s potential outcome to depend \emph{both} on the treatment she actually receives, $D_i$, and the experimental arm to which she is assigned, $Z_i$. This model is too general, however, to allow us to explore the importance of paternalism. Accordingly, we impose the following standard exclusion restriction:
\begin{equation}
Y_i(d=0, z=0) = Y_i(d=0, z=2) \equiv Y_{i0}, \quad
Y_i(d=1, z=1) = Y_i(d=1, z=2) \equiv Y_{i1}.
\label{eq:exclusion}
\end{equation}
\ref{eq:exclusion} says that each participant's potential outcome is the same regardless of whether she \emph{chooses} a particular treatment $D=d$ or is \emph{assigned} the same treatment. Viewing $Z_i$ as an instrumental variable (IV) for the treatment $D_i$, this nothing more than the usual IV exclusion restriction.\footnote{\cite{chamberlain2011bayesian} uses an identical assumption to develop a theory of optimal treatment choice for an individual who has access to data from an RCT.} \ref{eq:exclusion} is plausible in our empirical setting and has testable implications that we fail to reject: see Appendix \todo{Fill in}{???} for details. Under \ref{eq:exclusion}, each participant's observed outcome $Y_i$ is related to her potential outcomes $(Y_{i0}, Y_{i1})$ according to 
\todo[inline]{Issac: add a section \ref{bound_tot_tut_appendix} to the appendix and drop in the results for the Huber \& Mellace stuff; Frank: follow-up and draft that section of the appendix.}
\begin{equation}
    Y_i = \mathbbm{1}(Z_i =0) Y_{i0} + \mathbbm{1}(Z_i = 1)  Y_{i1}  + \mathbbm{1}(Z_i = 2) \left[(1 - C_i) Y_{i0} + C_i Y_{i1} \right].
\label{eq:potentialOutcomes}
\end{equation}

\ref{eq:potentialOutcomes} is the key to understanding the results that follows. By randomly assigning $Z_i=0$ and $Z_i = 1$, our experiment identifies the marginal distributions of $Y_{i0}$ and $Y_{i1}$ in the population as a whole. By randomly assigning $Z_i=2$, it likewise identifies the share of people with $C_i = 1$ (the treated), the distribution of $Y_{i1}$ for these individuals, and the distribution of $Y_{i0}$ for those with $C_i = 0$ (the untreated). Because $Z_i$ is assigned independently of pre-treatment covariates $X_i$, our design also identifies the corresponding \emph{conditional} distributions of $Y_{i0}$ and $Y_{i1}$ given $X_i$. These distributions are the ingredients that we use below to explore the importance of paternalism.

@@

In the second step we fit a linear regression model to data from the training set using regressors constructed from the proxy functions $\psi(\cdot|\text{Training})$ and $\tau(\cdot|\text{Training})$ constructed in the first step. In particular, we estimate 
\begin{equation}
Y_i = \alpha_0 + \alpha_1 \psi_i + \beta_1 (Z_i - \mathbbm{E}[Z_i]) + \beta_2 (Z_i - \mathbbm{E}[Z_i])(\tau_i - \mathbbm{E}[\tau_i]) + \epsilon_i
\label{eq:chernreg}
\end{equation}
where $\psi_i \equiv \psi(X_i|\text{Training})$ and $\tau_i \equiv \tau(X_i|\text{Training})$.\footnote{This is a slightly simpler regression than the one proposed in equation (3.1) of \cite{chernozhukov2018generic}, which involves propensity score weights. Because the random assignment of $Z$ in our experiment does \emph{not} condition on $X$, the propensity score weights in our case are constant over $X$ and hence drop out.} As shown by \cite{chernozhukov2018generic}, the coefficients $\beta_1$ and $\beta_2$ from \ref{eq:chernreg} identify the \emph{best linear predictor} of the conditional ATE based on $\tau(\cdot|\text{Training})$, namely
\[
\beta_1 = \mathbbm{E}[\text{ATE}(X_i)] = \text{ATE}, \quad
\beta_2 = \frac{\text{Cov}[\text{ATE}(X_i), \tau_i]}{\text{Var}(\tau_i)}.
\]
If treatment effects are homogeneous we must have $\beta_2 = 0$. Rejecting this hypothesis establishes that $\tau_i$ predicts $\text{ATE}(X_i)$ and hence that $\Delta_i$ varies. Since $\tau_i$ and $\psi_i$ do not depend on the test set, inference for the regression in \ref{eq:chernreg} is straightforward conditional on the Training/Test split.

@@

The key insight can be read directly from \ref{eq:potentialTreatments} and \ref{eq:potentialOutcomes}. Viewing $Z_i$ as an instrumental variable, our design can be interpreted as a \emph{pair} of RCTs, each subject to one-sided non-compliance.\footnote{For a full derivation, see \hl{Appendix ???}.} The first of these compares $Z_i=0$ to $Z_i = 2$. For each individual with $Z_i = 0$ we have $D_i = 0$ and observe $Y_{i0}$. For those with $Z_i = 2$ we have $D_i = C_i$ and observe $(1 - C_i) Y_{i0} + C_i Y_{i1}$. This is identical to a ``randomized encouragement'' design in which treatment is only available to those who are encouraged: $Z_i = 2$. Under this interpretation, those with $C_i = 1$ are ``the compliers'' and it follows that 
\begin{equation}
\frac{\mathbbm{E}(Y_i|Z_i=2) - \mathbbm{E}(Y_i|Z_i =0)}{\mathbbm{E}(D_i|Z_i=2)-\mathbbm{E}(D_i|Z_i=0)} = 
\frac{\mathbbm{E}(Y_i|Z_i=2) - \mathbbm{E}(Y_i|Z_i =0)}{\mathbbm{E}(D_i|Z_i=2)} = \mathbbm{E}(Y_{i1} - Y_{i0}|C_i = 1)
\label{eq:ToT}
\end{equation}
since $\mathbbm{E}(D_i|Z_i=0)=0$ by \ref{eq:potentialTreatments}.

A closely related if slightly less intuitive argument can be used to construct a Wald estimand that identifies the TuT. Here we consider $Z_i = 1$ to be the ``encouragement'' and compare the outcomes for these individuals to those with $Z_i = 2$. If $Z_i = 1$ then $D_i = 1$ and we observe $Y_{i1}$; If $Z_i = 2$ then $D_i = C_i$ and we observe $(1 - C_i) Y_{i0} + C_i Y_{i1}$. Again, we can view this as an experiment with one-sided non-compliance, but now the situation is reversed. Everyone with $Z_i = 1$ is treated, but some people with $Z_i = 2$ are ``always-takers'' who obtain the treatment ($D_i = 1$) despite having been allocated to the ``control'' arm $Z_i=2$. Under this interpretation, the ``compliers'' are those with $C_i = 0$: when $Z_i=1$ they take the treatment, and when $Z_i=2$, they do not. Thus, 
\begin{equation}
\frac{\mathbbm{E}(Y_i|Z_i=1) - \mathbbm{E}(Y_i|Z_i =2)}{\mathbbm{E}(D_i|Z_i=1)-\mathbbm{E}(D_i|Z_i=2)} = 
\frac{\mathbbm{E}(Y_i|Z_i=1) - \mathbbm{E}(Y_i|Z_i =0)}{1 - \mathbbm{E}(D_i|Z_i=2)} = \mathbbm{E}(Y_{i1} - Y_{i0} | C_i = 0)
\label{eq:TuT}
\end{equation}
since $\mathbbm{E}(D_i|Z_i = 1)= 1$ by \ref{eq:potentialTreatments}.

\ref{eq:ToT} and \ref{eq:TuT} are useful for understanding why our design identifies the ToT and TuT, but they are less convenient for estimation and inference. In Appendix \todo{Frank: fill in}{???}, we show that
\begin{align}
\label{eq:ToTreg}
Y_i &= \mathbbm{E}(Y_{i0}) + (\text{ATE}) \mathbbm{1}(Z_i = 1) + (\text{ToT}) \left[\mathbbm{1}(Z_i = 2) \times D_i\right] + U_i \\
\label{eq:TuTreg}
Y_i &= \mathbbm{E}(Y_{i1}) + (\text{ATE}) \left[ -\mathbbm{1}(Z_i = 0)\right] + (\text{TuT}) \left[ -\mathbbm{1}(Z_i = 2) \times (1 - D_i)\right] + V_i 
\end{align}
where the error terms $U_i$ and $V_i$ satisfy 
\[
\mathbbm{E}(V_i) = \mathbbm{E}(U_i) = \mathbbm{E}[U_i \mathbbm{1}(Z_i = z)] = \mathbbm{E}[V_i \mathbbm{1}(Z_i = z)] = 0, \quad z = 0, 1, 2.
\]
It follows that a pair of just-identified, linear instrumental variables regressions can be used to estimate and carry out inference for each of our quantities of interest. To identify the ATE and ToT, run IV on \ref{eq:ToTreg} with instruments $\mathbbm{1}(Z_i = 1)$ and $\mathbbm{1}(Z_i=2)$. An F-test based on this regression can then be used to test the restriction $\text{ATE} = \text{ToT}$, and the usual IV output can be used to carry out inference for the ToT. No programming is required. Similarly, to identify the ATE and TuT run IV on \ref{eq:TuTreg} with instruments $\mathbbm{1}(Z_i = 0)$ and $\mathbbm{1}(Z_i = 2)$. An F-test based on this regression can then be used to test the restriction $\text{ATE} = \text{TuT}$, and the usual IV output can be used to carry out inference for the TuT. Again, no programming or special software is required.

@@

%Define the shorthand
%\[
%p_0 \equiv \mathbbm{P}(Z_i=0), \quad p_1 \equiv \mathbbm{P}(Z_i = 1), \quad q_0 \equiv \mathbbm{E}[\mathbbm{1}(Z_i = 2)(1 - D_i)], \quad q_1 \equiv \mathbbm{E}[\mathbbm{1}(Z_i = 2) D_i].
%\]
%As shown in Appendix \todo{Frank: fill in}{???}, the asymptotic variance of $(\text{Tot} - \text{TuT})$ is given by
%\[
%\text{AVAR}(\text{ToT} - \text{TuT}) = \boldsymbol{\omega}' \boldsymbol{\Sigma} \boldsymbol{\omega}, \quad
%\boldsymbol{\omega}' \equiv \displaystyle \begin{bmatrix}
%-\frac{1}{q_1} & \frac{1}{q_0} & \left(\frac{p_1 - 1}{p_0q_1} - \frac{1}{q_0}\right) & -\frac{1}{q_1} - \frac{1 - p_0}{p_1 q_0}
%\end{bmatrix}
%\]
%where $\boldsymbol{\Sigma}$ is a covariance matrix that can be estimated from the residuals of \ref{eq:ToTreg} and \ref{eq:TuTreg}. Under iid observations, 
%
%\todo[inline]{Explain two leading forms that $\Sigma$ can take: heterosked and cluster}
